---
title: "GNS-I: Gradient Boosted Modeling, Feature Selection & Hyperparameter Tuning"
date: June 14, 2026
author:
  - name: Joseph G. Nugent, MD, MPH, MHS, CPH
    orcid: 0000-0002-6863-4523
    email: nugentj@ohsu.edu
execute:
  echo: true
  warning: false
format:
  html:
    anchor-sections: false
    cap-location: top
    citations-hover: true
    code-copy: true
    code-fold: true
    code-line-numbers: true
    code-overflow: wrap
    code-tools: true
    crossref:
      fig-title: '<strong>Figure'
      tbl-title: '<strong>Table'
      title-delim: '</strong>:'
    css: /Users/JosephNugent/GitHub/quarto/_includes/styles.css
    df-print: paged
    embed-resources: true
    fig-align: center
    fig-width: 8
    fontsize: 17px
    html-math-method: katex
    link-external-icon: true
    link-external-newwindow: true
    mainfont: EB Garamond
    number-sections: false
    page-layout: full
    theme:
      - default
      - /Users/JosephNugent/GitHub/quarto/_includes/custom.scss
    toc: true
    toc-title: Contents
---

```{r}
#| label: setup
#| include: false

library(easystats)
librarian::shelf(
  finetune,
  jgntools,
  tidymodels,
  tidyverse,
  xgboost
)
```

```{r}
#| label: Define functions

theme_set(theme_jgn())

# Disable scientific notation
options(scipen = 999)
options(knitr.table.format = 'html')
#options(knitr.table.format = 'pipe')
options(knitr.kable.NA = '')

jgntools::resolve_conflicts()

sysfonts::font_add_google('EB Garamond')

font <- 'EB Garamond'

jgn_xgboost_cleanup <- function(prefix = NA){
  if(is.na(prefix)){
  	stop('`prefix` is required to proceed.')
  }

	variables <- c(
	  'outcome',
    'outcome_prevalence',
    # 'cv_folds_display',
	  'xgboost_best_overall',
    'xgboost_data',
    'xgboost_data_split',
    'xgboost_data_test',
    'xgboost_data_train',
    'xgboost_final_model',
    'xgboost_final_recipe',
    'xgboost_final_workflow_production',
    'xgboost_grid',
    'xgboost_last_fit',
    'xgboost_metrics',
    'xgboost_nested_results',
    'xgboost_outer_folds',
    'xgboost_outer_metrics',
    'xgboost_predictions',
    'xgboost_recipe',
    'xgboost_roc',
    'xgboost_roc_coords',
    'xgboost_roc_coords_ci',
    'xgboost_shap',
    'xgboost_shap_x',
    'xgboost_spec',
    'xgboost_workflow'
  )

  if(exists('gamma_parameter')){
    variables <- c(
      'gamma_parameter',
      variables
    )
  }

  if(exists('num_trees')){
    variables <- c(
      'num_trees',
      variables
    )
  }

  walk(
    variables,
    ~assign(paste0(prefix, '_', .x), get(.x), get(.x), envir = .GlobalEnv)
  )

  rm(list = variables, envir = .GlobalEnv)
}
```

```{r}
#| label: Load Stored Dataset Objects

load('./GNS-I_Analytic Dataset.RData')
```

# Outcome `xgboost` Hyperparameter Tuning

::: {.panel-tabset}

## Lasting Medical Complications

::: {.panel-tabset}

```{r}
#| label: lasting_complications specification

xgboost_data <- gns %>%
  filter(!is.na(lasting_complications))

set.seed(42)
xgboost_data_split <- initial_split(
  xgboost_data,
  strata = lasting_complications
)
xgboost_data_train <- training(xgboost_data_split)
xgboost_data_test <- testing(xgboost_data_split)

xgboost_outer_folds <- vfold_cv(
  xgboost_data_train,
  v = 5,
  repeats = 1,
  strata = lasting_complications
)

xgboost_recipe <-
  recipe(formula = lasting_complications ~ ., data = xgboost_data_train) %>%
  step_rm(
    one_of(
      'institution',
      # Outcomes
      'complications_fever', 'complications_infection', 'complications_spasticity', 'complications_coagulopathy', 'complications_wound', 'complications_csf_leak', 'complications_subdural_hygroma', 'complications_nutritional', 'complications_other', 'infection_location_unknown', 'infection_location_urt', 'infection_location_lrt', 'infection_location_gu', 'infection_location_gi', 'infection_location_eye', 'infection_location_skin', 'infection_location_perinephric', 'infection_location_blood', 'infection_location_other', 'bleeding_diathesis', 'wound_malalignment', 'wound_stitch_sinus', 'wound_fluid_collection',
      # 'lasting_complications',
      'mortality_discharge', 'mortality_24h', 'mortality_30d', 'mortality_90d',
      'seizures_longterm', 'shunt_longterm',
      'gos_30d', 'gos_90d',
      'discharge_destination',
      # Extraneous data
      'gini_year',
      # Character variables
      'record_id', 'country_code', 'sedation_other_meds', 'analgesia_other_agents', 'analgesia_max_dose', 'intraoperative_adverse_events_list', 'reoperation_other_indication', 'icp_meds_other_agent', 'hyperosm_other', 'hyperosm_dose', 'feeding_complications'
    ),
    ends_with('_complete')
  ) %>%
  #update_role(record_id, new_role = 'ID') %>%
  step_date(where(is.Date) | where(is.POSIXct), features = c('dow', 'doy', 'week', 'month', 'quarter', 'semester', 'year')) %>%
  step_time(where(is.POSIXlt) | where(is.POSIXct)) %>%
  step_mutate_at(where(is.POSIXct), fn = as.POSIXct.Date) %>%
  step_holiday(where(is.Date), holidays = timeDate::listHolidays()) %>%
  step_rm(where(is.Date) | where(is.POSIXlt) | where(is.POSIXct)) %>%
  step_novel(all_nominal_predictors()) %>%
  step_unknown(all_nominal_predictors(), new_level = 'Missing') %>%
  step_rm(ends_with('_Missing')) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_zv(all_predictors()) %>%
  step_nzv(all_predictors())
```

```{r}
#| label: lasting_complications xgboost

xgboost_spec <-
  boost_tree(
    trees        = tune(),
    tree_depth   = tune(),
    min_n        = tune(),
    learn_rate   = tune(),
    loss_reduction = 20, #tune(),
    sample_size  = tune(),
    mtry         = tune()
  ) %>%
  set_engine('xgboost') %>%
  set_mode('classification')

xgboost_workflow <-
  workflow() %>%
  add_recipe(xgboost_recipe) %>%
  add_model(xgboost_spec)

xgboost_grid <-
  grid_space_filling(
    trees(range = c(100L, 1000L)),
    tree_depth(),
    min_n(),
    learn_rate(range = c(-3, -1)),   # log10 scale → 0.001–0.1
    # loss_reduction(),
    sample_prop(),                    # maps to sample_size in boost_tree
    finalize(mtry(), xgboost_data_train),          # resolves upper bound from data
    size = 40
  )

xgboost_nested_results <- xgboost_outer_folds$splits %>%
  purrr::imap(\(xgboost_outer_split, fold_id) {

    cat('--- Outer fold', fold_id, '---\n')

    # Inner resampling on the outer *analysis* set
    xgboost_inner_folds <- vfold_cv(
      analysis(xgboost_outer_split),
      v       = 4,
      strata  = lasting_complications
    )

    # Tune on inner folds (race ANOVA drops poor candidates early)
    xgboost_tune <- tune_race_anova(
      xgboost_workflow,
      resamples = xgboost_inner_folds,
      grid      = xgboost_grid,
      metrics   = metric_set(roc_auc, pr_auc, accuracy, bal_accuracy, kap, mcc, mn_log_loss),
      control   = control_race(
        verbose = FALSE,
        save_pred = FALSE,
        event_level = 'second'
      )
    )

    # Best hyperparameters from inner CV
    xgboost_best_params <- select_best(xgboost_tune, metric = 'roc_auc')

    # Finalize workflow and evaluate on the outer *assessment* set
    xgboost_finalized_workflow <- finalize_workflow(xgboost_workflow, xgboost_best_params)

    xgboost_outer_fit <- last_fit(xgboost_finalized_workflow, split = xgboost_outer_split)

    # Return tidy results
    list(
      fold         = fold_id,
      xgboost_best_params  = xgboost_best_params,
      metrics      = collect_metrics(xgboost_outer_fit),
      predictions  = collect_predictions(xgboost_outer_fit)
    )
  })

xgboost_outer_metrics <-
  xgboost_nested_results %>%
  purrr::map(\(r) mutate(r$metrics, fold = r$fold)) %>%
  purrr::list_rbind()

# xgboost_outer_metrics %>%
#   summarise(
#     mean_roc_auc = mean(.estimate[.metric == 'roc_auc']),
#     sd_roc_auc   = sd(.estimate[.metric == 'roc_auc']),
#     mean_accuracy = mean(.estimate[.metric == 'accuracy']),
#     .by = .metric
#   )

xgboost_best_overall <-
  xgboost_nested_results %>%
  purrr::map(\(r) {
    r$xgboost_best_params %>% mutate(roc_auc = r$metrics$.estimate[r$metrics$.metric == 'roc_auc'][1])
  }) %>%
  purrr::list_rbind() %>%
  slice_max(roc_auc, n = 1)

xgboost_final_workflow_production <- finalize_workflow(
  xgboost_workflow,
  xgboost_best_overall
)

xgboost_last_fit <- last_fit(
  xgboost_final_workflow_production,
  split = xgboost_data_split
)

xgboost_metrics <- xgboost_last_fit %>%
  collect_metrics()

xgboost_predictions <- xgboost_last_fit %>%
  collect_predictions()
```

### Summary Table

```{r}
xgboost_best_overall %>% 
  mutate(
    `Inner Cross-Validation Folds` = 3,
    `Outer Cross-Validation Folds` = 5,
    `Minimum Loss Reduction` = 20
  ) %>% 
  rename(
    `Trees` = trees,
    `Maximum Tree Depth` = tree_depth,
    `Minimal Node Size` = min_n,
		`Proportion Observations Sampled` = sample_size,
		`Randomly Selected Predictors` = mtry,
		`Learning Rate` = learn_rate
  ) %>% 
  select(
    `Inner Cross-Validation Folds`,
    `Outer Cross-Validation Folds`,
    `Minimum Loss Reduction`,
    `Trees`,
    `Maximum Tree Depth`,
    `Minimal Node Size`,
		`Proportion Observations Sampled`,
		`Randomly Selected Predictors`,
		`Learning Rate`
  ) %>% 
  print_kable()
```

### Confusion Matrix

```{r}
jgn_xgboost_confusion_matrix(outcome_variable = 'lasting_complications', positive_level_num = 2)
```

### AUROC

```{r}
xgboost_roc <- roc(
  response = xgboost_predictions %>% pull(lasting_complications),
  predictor = xgboost_predictions %>% pull(`.pred_Complication Occurrence`),
  ci = TRUE
)

xgboost_roc %>% 
  jggroc()
```

### AUPRC

```{r}
jgn_xgboost_prc(outcome_variable = 'lasting_complications', positive_label = 'Complication Occurrence', positive_level_label = 'second')
```

### Calibration Plot

```{r}
xgboost_predictions %>% 
  cal_plot_windowed(
    truth = lasting_complications,
    estimate = `.pred_Complication Occurrence`,
    conf_level = 0.95,
    event_level = 'second'
  ) + 
  theme_jgn() +
  labs(
    x = 'Predicted Probability',
    y = 'Empirical Probability'
  )
```

### Shapley Variable Importance

```{r}
xgboost_shap <- 
  xgboost_last_fit %>% 
  extract_fit_parsnip() %>% 
  extract_fit_engine()

xgboost_final_model <- xgboost_last_fit %>% 
  extract_fit_parsnip() %>% 
  extract_fit_engine()

xgboost_final_recipe <- xgboost_last_fit %>% 
  extract_recipe()

xgboost_shap_x <- xgboost_data_train %>% 
  bake(object = xgboost_final_recipe, new_data = ., all_predictors()) %>% 
  as.matrix()

xgboost_shap <- shapviz(xgboost_final_model, X_pred = xgboost_shap_x)

xgboost_shap$S %>% 
  as_tibble() %>% 
  summarise(
    across(
      everything(),
      ~mean(abs(.x))
    )
  ) %>% 
  pivot_longer(
    everything(),
    names_to = 'Variable',
    values_to = 'Importance'
  ) %>% 
  arrange(desc(Importance)) %>% 
  mutate(
    Sign = if_else(
      map_dbl(
        Variable,
        ~mean(xgboost_shap$S[, .x])
      ) >= 0, 'Positive', 'Negative'
    ) %>% factor()
  ) %>% 
  filter(Importance != 0) %>% 
  print_kable()
```

### Shapley Importance Overlay

```{r}
sv_importance(xgboost_shap, kind = 'both', show_numbers = TRUE, alpha = 0.2, width = 0.2) + 
  theme_bw(base_family = font) +
  theme(text = element_text(family = font, size = 14))
```

:::

```{r}
#| label: Cleanup lasting_complications

jgn_xgboost_cleanup('lasting_complications')
```

## Mortality at Discharge

::: {.panel-tabset}

```{r}
#| label: mortality_discharge specification

xgboost_data <- gns %>%
  filter(!is.na(mortality_discharge))

set.seed(42)
xgboost_data_split <- initial_split(
  xgboost_data,
  strata = mortality_discharge
)
xgboost_data_train <- training(xgboost_data_split)
xgboost_data_test <- testing(xgboost_data_split)

xgboost_outer_folds <- vfold_cv(
  xgboost_data_train,
  v = 5,
  repeats = 1,
  strata = mortality_discharge
)

xgboost_recipe <-
  recipe(formula = mortality_discharge ~ ., data = xgboost_data_train) %>%
  step_rm(
    one_of(
      #'admission_interval',
      'redcap_event_name',
      'redcap_data_access_group',
      'length_of_stay_redcap',
      'institution',
      # Outcomes
      'lasting_complications',
      #'mortality_discharge',
      'mortality_24h', 'mortality_30d', 'mortality_90d',
      'seizures_longterm', 'shunt_longterm',
      'gos_30d', 'gos_90d',
      'discharge_destination',
      # Extraneous data
      'gini_year',
      # Character variables
      'record_id', 'country_code', 'hosp_id', 'sedation_other_meds', 'analgesia_other_agents', 'analgesia_max_dose', 'intraoperative_adverse_events_list', 'reoperation_other_indication', 'icp_meds_other_agent', 'hyperosm_other', 'hyperosm_dose', 'feeding_complications'
    ),
    ends_with('_complete')
  ) %>%
  #update_role(record_id, new_role = 'ID') %>%
  step_date(where(is.Date) | where(is.POSIXct), features = c('dow', 'doy', 'week', 'month', 'quarter', 'semester', 'year')) %>%
  step_time(where(is.POSIXlt) | where(is.POSIXct)) %>%
  step_mutate_at(where(is.POSIXct), fn = as.POSIXct.Date) %>%
  step_holiday(where(is.Date), holidays = timeDate::listHolidays()) %>%
  step_rm(where(is.Date) | where(is.POSIXlt) | where(is.POSIXct)) %>%
  step_novel(all_nominal_predictors()) %>%
  step_unknown(all_nominal_predictors(), new_level = 'Missing') %>%
  step_rm(ends_with('_Missing')) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_zv(all_predictors()) %>%
  step_nzv(all_predictors())
```

```{r}
#| label: mortality_discharge xgboost

xgboost_spec <-
  boost_tree(
    trees        = tune(),
    tree_depth   = tune(),
    min_n        = tune(),
    learn_rate   = tune(),
    loss_reduction = 15, #tune(),
    sample_size  = tune(),
    mtry         = tune()
  ) %>%
  set_engine('xgboost') %>%
  set_mode('classification')

xgboost_workflow <-
  workflow() %>%
  add_recipe(xgboost_recipe) %>%
  add_model(xgboost_spec)

xgboost_grid <-
  grid_space_filling(
    trees(range = c(100L, 1000L)),
    tree_depth(),
    min_n(),
    learn_rate(range = c(-3, -1)),   # log10 scale → 0.001–0.1
    # loss_reduction(),
    sample_prop(),                    # maps to sample_size in boost_tree
    finalize(mtry(), xgboost_data_train),          # resolves upper bound from data
    size = 40
  )

xgboost_nested_results <- xgboost_outer_folds$splits %>%
  purrr::imap(\(xgboost_outer_split, fold_id) {

    cat('--- Outer fold', fold_id, '---\n')

    # Inner resampling on the outer *analysis* set
    xgboost_inner_folds <- vfold_cv(
      analysis(xgboost_outer_split),
      v       = 4,
      strata  = mortality_discharge
    )

    # Tune on inner folds (race ANOVA drops poor candidates early)
    xgboost_tune <- tune_race_anova(
      xgboost_workflow,
      resamples = xgboost_inner_folds,
      grid      = xgboost_grid,
      metrics   = metric_set(roc_auc, pr_auc, accuracy, bal_accuracy, kap, mcc, mn_log_loss),
      control   = control_race(
        verbose = FALSE,
        save_pred = FALSE,
        event_level = 'second'
      )
    )

    # Best hyperparameters from inner CV
    xgboost_best_params <- select_best(xgboost_tune, metric = 'roc_auc')

    # Finalize workflow and evaluate on the outer *assessment* set
    xgboost_finalized_workflow <- finalize_workflow(xgboost_workflow, xgboost_best_params)

    xgboost_outer_fit <- last_fit(xgboost_finalized_workflow, split = xgboost_outer_split)

    # Return tidy results
    list(
      fold         = fold_id,
      xgboost_best_params  = xgboost_best_params,
      metrics      = collect_metrics(xgboost_outer_fit),
      predictions  = collect_predictions(xgboost_outer_fit)
    )
  })

xgboost_outer_metrics <-
  xgboost_nested_results %>%
  purrr::map(\(r) mutate(r$metrics, fold = r$fold)) %>%
  purrr::list_rbind()

# xgboost_outer_metrics %>%
#   summarise(
#     mean_roc_auc = mean(.estimate[.metric == 'roc_auc']),
#     sd_roc_auc   = sd(.estimate[.metric == 'roc_auc']),
#     mean_accuracy = mean(.estimate[.metric == 'accuracy']),
#     .by = .metric
#   )

xgboost_best_overall <-
  xgboost_nested_results %>%
  purrr::map(\(r) {
    r$xgboost_best_params %>% mutate(roc_auc = r$metrics$.estimate[r$metrics$.metric == 'roc_auc'][1])
  }) %>%
  purrr::list_rbind() %>%
  slice_max(roc_auc, n = 1)

xgboost_final_workflow_production <- finalize_workflow(
  xgboost_workflow,
  xgboost_best_overall
)

xgboost_last_fit <- last_fit(
  xgboost_final_workflow_production,
  split = xgboost_data_split
)

xgboost_metrics <- xgboost_last_fit %>%
  collect_metrics()

xgboost_predictions <- xgboost_last_fit %>%
  collect_predictions()
```

### Summary Table

```{r}
xgboost_best_overall %>% 
  mutate(
    `Inner Cross-Validation Folds` = 3,
    `Outer Cross-Validation Folds` = 5,
    `Minimum Loss Reduction` = 15
  ) %>% 
  rename(
    `Trees` = trees,
    `Maximum Tree Depth` = tree_depth,
    `Minimal Node Size` = min_n,
		`Proportion Observations Sampled` = sample_size,
		`Randomly Selected Predictors` = mtry,
		`Learning Rate` = learn_rate
  ) %>% 
  select(
    `Inner Cross-Validation Folds`,
    `Outer Cross-Validation Folds`,
    `Minimum Loss Reduction`,
    `Trees`,
    `Maximum Tree Depth`,
    `Minimal Node Size`,
		`Proportion Observations Sampled`,
		`Randomly Selected Predictors`,
		`Learning Rate`
  ) %>% 
  print_kable()
```

### Confusion Matrix

```{r}
jgn_xgboost_confusion_matrix(outcome_variable = 'mortality_discharge', positive_level_num = 2)
```

### AUROC

```{r}
xgboost_roc <- roc(
  response = xgboost_predictions %>% pull(mortality_discharge),
  predictor = xgboost_predictions %>% pull(`.pred_Deceased`),
  ci = TRUE
)

xgboost_roc %>% 
  jggroc()
```

### AUPRC

```{r}
jgn_xgboost_prc(outcome_variable = 'mortality_discharge', positive_label = 'Deceased', positive_level_label = 'second')
```

### Calibration Plot

```{r}
xgboost_predictions %>% 
  cal_plot_windowed(
    truth = mortality_discharge,
    estimate = `.pred_Deceased`,
    conf_level = 0.95,
    event_level = 'second'
  ) + 
  theme_jgn() +
  labs(
    x = 'Predicted Probability',
    y = 'Empirical Probability'
  )
```

### Shapley Variable Importance

```{r}
xgboost_shap <- 
  xgboost_last_fit %>% 
  extract_fit_parsnip() %>% 
  extract_fit_engine()

xgboost_final_model <- xgboost_last_fit %>% 
  extract_fit_parsnip() %>% 
  extract_fit_engine()

xgboost_final_recipe <- xgboost_last_fit %>% 
  extract_recipe()

xgboost_shap_x <- xgboost_data_train %>% 
  bake(object = xgboost_final_recipe, new_data = ., all_predictors()) %>% 
  as.matrix()

xgboost_shap <- shapviz(xgboost_final_model, X_pred = xgboost_shap_x)

xgboost_shap$S %>% 
  as_tibble() %>% 
  summarise(
    across(
      everything(),
      ~mean(abs(.x))
    )
  ) %>% 
  pivot_longer(
    everything(),
    names_to = 'Variable',
    values_to = 'Importance'
  ) %>% 
  arrange(desc(Importance)) %>% 
  mutate(
    Sign = if_else(
      map_dbl(
        Variable,
        ~mean(xgboost_shap$S[, .x])
      ) >= 0, 'Positive', 'Negative'
    ) %>% factor()
  ) %>% 
  filter(Importance != 0) %>% 
  print_kable()
```

### Shapley Importance Overlay

```{r}
sv_importance(xgboost_shap, kind = 'both', show_numbers = TRUE, alpha = 0.2, width = 0.2) + 
  theme_bw(base_family = font) +
  theme(text = element_text(family = font, size = 14))
```

:::

```{r}
#| label: Cleanup mortality_discharge

jgn_xgboost_cleanup('mortality_discharge')
```

## Mortality at 90 days Post-Injury

::: {.panel-tabset}

```{r}
#| label: mortality_90d specification

xgboost_data <- gns %>%
  filter(!is.na(mortality_90d))

set.seed(42)
xgboost_data_split <- initial_split(
  xgboost_data,
  strata = mortality_90d
)
xgboost_data_train <- training(xgboost_data_split)
xgboost_data_test <- testing(xgboost_data_split)

xgboost_outer_folds <- vfold_cv(
  xgboost_data_train,
  v = 5,
  repeats = 1,
  strata = mortality_90d
)

xgboost_recipe <-
  recipe(formula = mortality_90d ~ ., data = xgboost_data_train) %>%
  step_rm(
    one_of(
      #'admission_interval',
      'redcap_event_name',
      'redcap_data_access_group',
      'length_of_stay_redcap',
      'institution',
      # Outcomes
      'lasting_complications',
      'mortality_discharge', 'mortality_24h', 'mortality_30d',
      #'mortality_90d',
      'seizures_longterm', 'shunt_longterm',
      'gos_30d', 'gos_90d',
      'discharge_destination',
      # Extraneous data
      'gini_year',
      # Character variables
      'record_id', 'country_code', 'hosp_id', 'sedation_other_meds', 'analgesia_other_agents', 'analgesia_max_dose', 'intraoperative_adverse_events_list', 'reoperation_other_indication', 'icp_meds_other_agent', 'hyperosm_other', 'hyperosm_dose', 'feeding_complications'
    ),
    ends_with('_complete')
  ) %>%
  #update_role(record_id, new_role = 'ID') %>%
  step_date(where(is.Date) | where(is.POSIXct), features = c('dow', 'doy', 'week', 'month', 'quarter', 'semester', 'year')) %>%
  step_time(where(is.POSIXlt) | where(is.POSIXct)) %>%
  step_mutate_at(where(is.POSIXct), fn = as.POSIXct.Date) %>%
  step_holiday(where(is.Date), holidays = timeDate::listHolidays()) %>%
  step_rm(where(is.Date) | where(is.POSIXlt) | where(is.POSIXct)) %>%
  step_novel(all_nominal_predictors()) %>%
  step_unknown(all_nominal_predictors(), new_level = 'Missing') %>%
  step_rm(ends_with('_Missing')) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_zv(all_predictors()) %>%
  step_nzv(all_predictors())
```

```{r}
#| label: mortality_90d xgboost

xgboost_spec <-
  boost_tree(
    trees        = tune(),
    tree_depth   = tune(),
    min_n        = tune(),
    learn_rate   = tune(),
    loss_reduction = 15, #tune(),
    sample_size  = tune(),
    mtry         = tune()
  ) %>%
  set_engine('xgboost') %>%
  set_mode('classification')

xgboost_workflow <-
  workflow() %>%
  add_recipe(xgboost_recipe) %>%
  add_model(xgboost_spec)

xgboost_grid <-
  grid_space_filling(
    trees(range = c(100L, 1000L)),
    tree_depth(),
    min_n(),
    learn_rate(range = c(-3, -1)),   # log10 scale → 0.001–0.1
    # loss_reduction(),
    sample_prop(),                    # maps to sample_size in boost_tree
    finalize(mtry(), xgboost_data_train),          # resolves upper bound from data
    size = 40
  )

xgboost_nested_results <- xgboost_outer_folds$splits %>%
  purrr::imap(\(xgboost_outer_split, fold_id) {

    cat('--- Outer fold', fold_id, '---\n')

    # Inner resampling on the outer *analysis* set
    xgboost_inner_folds <- vfold_cv(
      analysis(xgboost_outer_split),
      v       = 4,
      strata  = mortality_90d
    )

    # Tune on inner folds (race ANOVA drops poor candidates early)
    xgboost_tune <- tune_race_anova(
      xgboost_workflow,
      resamples = xgboost_inner_folds,
      grid      = xgboost_grid,
      metrics   = metric_set(roc_auc, pr_auc, accuracy, bal_accuracy, kap, mcc, mn_log_loss),
      control   = control_race(
        verbose = FALSE,
        save_pred = FALSE,
        event_level = 'second'
      )
    )

    # Best hyperparameters from inner CV
    xgboost_best_params <- select_best(xgboost_tune, metric = 'roc_auc')

    # Finalize workflow and evaluate on the outer *assessment* set
    xgboost_finalized_workflow <- finalize_workflow(xgboost_workflow, xgboost_best_params)

    xgboost_outer_fit <- last_fit(xgboost_finalized_workflow, split = xgboost_outer_split)

    # Return tidy results
    list(
      fold         = fold_id,
      xgboost_best_params  = xgboost_best_params,
      metrics      = collect_metrics(xgboost_outer_fit),
      predictions  = collect_predictions(xgboost_outer_fit)
    )
  })

xgboost_outer_metrics <-
  xgboost_nested_results %>%
  purrr::map(\(r) mutate(r$metrics, fold = r$fold)) %>%
  purrr::list_rbind()

# xgboost_outer_metrics %>%
#   summarise(
#     mean_roc_auc = mean(.estimate[.metric == 'roc_auc']),
#     sd_roc_auc   = sd(.estimate[.metric == 'roc_auc']),
#     mean_accuracy = mean(.estimate[.metric == 'accuracy']),
#     .by = .metric
#   )

xgboost_best_overall <-
  xgboost_nested_results %>%
  purrr::map(\(r) {
    r$xgboost_best_params %>% mutate(roc_auc = r$metrics$.estimate[r$metrics$.metric == 'roc_auc'][1])
  }) %>%
  purrr::list_rbind() %>%
  slice_max(roc_auc, n = 1)

xgboost_final_workflow_production <- finalize_workflow(
  xgboost_workflow,
  xgboost_best_overall
)

xgboost_last_fit <- last_fit(
  xgboost_final_workflow_production,
  split = xgboost_data_split
)

xgboost_metrics <- xgboost_last_fit %>%
  collect_metrics()

xgboost_predictions <- xgboost_last_fit %>%
  collect_predictions()
```

### Summary Table

```{r}
xgboost_best_overall %>% 
  mutate(
    `Inner Cross-Validation Folds` = 3,
    `Outer Cross-Validation Folds` = 5,
    `Minimum Loss Reduction` = 20
  ) %>% 
  rename(
    `Trees` = trees,
    `Maximum Tree Depth` = tree_depth,
    `Minimal Node Size` = min_n,
		`Proportion Observations Sampled` = sample_size,
		`Randomly Selected Predictors` = mtry,
		`Learning Rate` = learn_rate
  ) %>% 
  select(
    `Inner Cross-Validation Folds`,
    `Outer Cross-Validation Folds`,
    `Minimum Loss Reduction`,
    `Trees`,
    `Maximum Tree Depth`,
    `Minimal Node Size`,
		`Proportion Observations Sampled`,
		`Randomly Selected Predictors`,
		`Learning Rate`
  ) %>% 
  print_kable()
```

### Confusion Matrix

```{r}
jgn_xgboost_confusion_matrix(outcome_variable = 'mortality_90d', positive_level_num = 2)
```

### AUROC

```{r}
xgboost_roc <- roc(
  response = xgboost_predictions %>% pull(mortality_90d),
  predictor = xgboost_predictions %>% pull(`.pred_Deceased`),
  ci = TRUE
)

xgboost_roc %>% 
  jggroc()
```

### AUPRC

```{r}
jgn_xgboost_prc(outcome_variable = 'mortality_90d', positive_label = 'Deceased', positive_level_label = 'second')
```

### Calibration Plot

```{r}
xgboost_predictions %>% 
  cal_plot_windowed(
    truth = mortality_90d,
    estimate = `.pred_Deceased`,
    conf_level = 0.95,
    event_level = 'second'
  ) + 
  theme_jgn() +
  labs(
    x = 'Predicted Probability',
    y = 'Empirical Probability'
  )
```

### Shapley Variable Importance

```{r}
xgboost_shap <- 
  xgboost_last_fit %>% 
  extract_fit_parsnip() %>% 
  extract_fit_engine()

xgboost_final_model <- xgboost_last_fit %>% 
  extract_fit_parsnip() %>% 
  extract_fit_engine()

xgboost_final_recipe <- xgboost_last_fit %>% 
  extract_recipe()

xgboost_shap_x <- xgboost_data_train %>% 
  bake(object = xgboost_final_recipe, new_data = ., all_predictors()) %>% 
  as.matrix()

xgboost_shap <- shapviz(xgboost_final_model, X_pred = xgboost_shap_x)

xgboost_shap$S %>% 
  as_tibble() %>% 
  summarise(
    across(
      everything(),
      ~mean(abs(.x))
    )
  ) %>% 
  pivot_longer(
    everything(),
    names_to = 'Variable',
    values_to = 'Importance'
  ) %>% 
  arrange(desc(Importance)) %>% 
  mutate(
    Sign = if_else(
      map_dbl(
        Variable,
        ~mean(xgboost_shap$S[, .x])
      ) >= 0, 'Positive', 'Negative'
    ) %>% factor()
  ) %>% 
  filter(Importance != 0) %>% 
  print_kable()
```

### Shapley Importance Overlay

```{r}
sv_importance(xgboost_shap, kind = 'both', show_numbers = TRUE, alpha = 0.2, width = 0.2) + 
  theme_bw(base_family = font) +
  theme(text = element_text(family = font, size = 14))
```

:::

```{r}
#| label: Cleanup mortality_90d

jgn_xgboost_cleanup('mortality_90d')
```

:::


***

:::{.callout-note .raw-code collapse="true" icon=false}

# `R` Session Information

```{r}
jgntools::print_session_info()
```

:::